---
title: "Effect of mutation types on expression levels"
author: "EXAMPLE"
date: '2022-06-15'
output: html_document
---

# Project: Effect of mutation types on expression levels 

## Load necessary packages

```{r}
library("tidyverse")
```

## load dataset into RStudio

```{r}
download.file("https://ndownloader.figshare.com/files/29148666", "data/mutation_types.csv")
d_data <- read.csv("data/mutation_types.csv")
```

## start analysis

## Use TP53 as a test case

```{r}
d_plot <- d_data %>% 
         select(contains("TP53")) %>% 
         filter(!TP53_substitution_category %in% c(NA, "splice-site")) %>% 
         drop_na() %>% 
         mutate("log10_TP53_expr" = log10(TP53_expr)) # explain to students purpose of transform

p <- ggplot(data = d_plot, mapping = aes(x = TP53_substitution_category, 
                                        y = log10_TP53_expr, 
                                        fill = TP53_substitution_category)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(color = "black", alpha = 0.5) +
    theme_bw() +
    theme(legend.position = "none")
```


## save plot to "plots" directory

Notes:
* students would have created the "plots" directory in class
* alternatively, create a different folder and change path below accordingly

```{r}
ggsave(filename = "plots/mutation_types.pdf",
       plot = p,
       width = 3, height = 3)
```

## Notes:
## - if students finish early, allow them to make plots for the other genes

