---
title: "Tumor Suppressors"
author: "EXAMPLE"
date: '2022-06-15'
output: html_document
---

# Project: Mutations in Tumor Suppressors 

## Load necessary packages

```{r}
library("tidyverse")
```

## load dataset into RStudio

Students would have created the "data" directory in class
Alternatively, create a different folder and change path below accordingly

```{r}
download.file("https://ndownloader.figshare.com/files/30366351", "data/tumor_suppressor.csv")
d_data <- read.csv("data/tumor_suppressor.csv") 
```

## start analysis

# create a histogram of the protein changes
If the students want, they could make a lollipop plot and also include a gene schematic using geom_shape

```{r}
d_plot <- d_data %>%
    separate_rows(TP53_protein_change, sep = ",") %>%  ## There's at least one with two mutations sep by a comma
    filter(!is.na(TP53_protein_change) & TP53_protein_change != "unknown") %>%
    select(TP53_protein_change) %>%
    mutate(TP53_aa_changed = parse_number(TP53_protein_change)) %>%
    ## Could generate colors based on type of mutations but I'll just use region of the protein here
    ## https://en.wikipedia.org/wiki/P53#/media/File:P53_Schematic.tif
    mutate(TP53_region = case_when( #mentor will have to teach case_when or provide this
      TP53_aa_changed <= 63 ~ "transactivation_domains",
      TP53_aa_changed <= 92 ~ "proline_rich_region",
      TP53_aa_changed <= 293 & TP53_aa_changed >= 102  ~ "dna_binding_domain",
      TP53_aa_changed <= 355 & TP53_aa_changed >= 320  ~ "tetramerization_domain",
      TP53_aa_changed >= 356  ~ "regulatory_domain",
      TRUE ~ "unspecified"
    ))

p <- ggplot(data = d_plot) +
  geom_histogram(aes(x = TP53_aa_changed, fill = TP53_region), binwidth = 5) +
  theme_bw()
```

## save plot to "plots" directory

Notes:
* students would have created the "plots" directory in class
* alternatively, create a different folder and change path below accordingly

```{r}
ggsave(filename = "plots/mutation_locations.pdf",
       plot = p,
       width = 7, height = 3)
```

