---
title: "Co-expression between genes"
author: "EXAMPLE"
date: '2022-06-15'
output: html_document
---

# Project: Co-expression between genes 

## Load necessary packages

```{r}
library("tidyverse")
```

## load dataset into RStudio

```{r}
download.file("https://ndownloader.figshare.com/files/29228802", "data/coexpression.csv")
d_data <- read.csv("data/coexpression.csv",
                  col_types = cols(TP53_substitution = col_character(),
                                   TP53_substitution_category = col_character()))
```

# load TP53 and RUNX target info

```{r}
download.file("https://ndownloader.figshare.com/files/29229288", "data/coexpression_supplement.csv")
d_targets <- read.csv("data/coexpression_supplement.csv")
```

## start analysis

Use RUNX2 as a test case
* largely works, positive correlation for a lot of the available cancers

```{r}
d_plot <- d_data %>% 
         select(project_id,  RUNX2_expr, TNFRSF19_expr) %>% 
         mutate(across(ends_with("expr"),
                       .fns = list(log10 = ~ log10(.))))
                       
p <- ggplot(data = d_plot, mapping = aes(x = RUNX2_expr_log10, 
                                        y = TNFRSF19_expr_log10)) +
    geom_point() +
    stat_smooth(method = "lm", col = "red") +
    theme_bw() +
    facet_wrap(vars(project_id))
```

Notes:
* students would have created the "plots" directory in class
* alternatively, create a different folder and change path below accordingly

```{r}
ggsave("plots/coexpression.pdf", p)
```
