---
title: "Cancer Immunology"
author: "EXAMPLE"
date: '2022-06-15'
output: html_document
---

# Project: Cancer immunology--tumor-infiltrating lymphocytes 

## Load necessary packages

```{r}
library("tidyverse")
```

## load dataset into RStudio
students have created the 'data' folder during the lessons
change path if new directory is created

```{r}
download.file("https://ndownloader.figshare.com/files/30290958", "data/survival.csv")
d_data <- read.csv("data/survival.csv") 
```

## start analysis

Example analysis 1 

Notes:
* students would have created the "plots" directory in class
* alternatively, create a different folder and change path below accordingly

```{r}
p_scatter <- d_data %>%
  filter(sample_type == "Primary Tumor") %>% 
  filter(project_id == "TCGA-LUAD") %>% 
  ggplot(aes(x=immune.score, y=B2M_expr)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method='lm', formula= y~x) +
  theme_classic()

ggsave("plots/scatter_immunescore_b2m.pdf", p_scatter)
```

Example analysis 2: Is granzyme b expression correlated with CD8 infiltration?

```{r}
p_scatter <- d_data %>%
  filter(sample_type == "Primary Tumor") %>%
  filter(project_id == "TCGA-LUAD") %>% 
  ggplot(aes(x=CD8.Tcell, y=GZMB_expr)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = 'lm', formula = y~x) +
  theme_classic() 

ggsave("plots/scatter_cd8_gzmb.pdf", p_scatter)
``` 


