---
title: "Patient Survival"
date: '2022-06-15'
output: html_document
---

# Project: Patient survival outcomes based on gene expression

## Load necessary packages

```{r}
library("tidyverse")
```

## load dataset into RStudio

Students have created the 'data' folder during the lessons
Change path if new directory is created

```{r}
download.file("https://ndownloader.figshare.com/files/29163735", "data/survival.csv")
d_data <- read.csv("data/survival.csv") 
```

## start analysis

Suggested workflow for students (feel free to change as you'd like)
Tasks:
- Explore the dataset
- Coordinate with your mentors, pick a cancer to work with
- Create a histogram to illustrate the distribution of EGFR expression for the patients in the cancer group you selected
- In your plot, highlight patients with high expression and low expression for EGFR
- Create a Kaplan-Meier plot to show cancer survival differences between high-vs-low EGFR expression patients

Example: create a histogram of EGFR_expr to illustrate high-vs-low values (using quartiles)

```{r}
d_plot <- d_data %>% 
         filter(project_id == "TCGA-BLCA") %>% 
         filter(sample_type == "Primary Tumor") %>% 
         mutate("log_EGFR_expr" = log10(EGFR_expr))

percentiles <- quantile(d_plot$log_EGFR_expr, c(0.25, 0.75))

p <- ggplot(d_plot, aes(x = log_EGFR_expr)) +
    geom_histogram(bins = 60, fill = "gray", color = "black") +
    theme_bw() +
    geom_vline(xintercept = percentiles[1], color = "red") +
    geom_vline(xintercept = percentiles[2], color = "red") +
    labs(x = "log10[EGFR expression (FPKM)]")
```

## save plot to "plots" directory

Notes:
* students would have created the "plots" directory in class
* alternatively, create a different folder and change path below accordingly

```{r}
ggsave(filename = "plots/histogram_TCGA_BLCA.pdf",
       plot = p,
       width = 3,
       height = 3)
```

Example 2--Create KM curves:

```{r}
# install packages
install.packages("survival")
install.packages("survminer")

# load packages
library("survival")
library("survminer")
```

```{r}
# BLCA works nicely: high EGFR expression correlated with poor survival outcomes
# stratification: top and bottom 25% or 33.3% both work
d_plot <- d_data %>% 
         filter(project_id == "TCGA-BLCA") %>% 
         filter(sample_type == "Primary Tumor")

fit <- survfit(Surv(time, status) ~ tercile_groups, data = d_plot)

p <- ggsurvplot(fit,
                pval = TRUE, 
                conf.int = TRUE,
                linetype = "strata", # Change line type by groups
                ggtheme = theme_classic(), # Change ggplot2 theme
                palette = c("#E7B800", "#2E9FDF"))
```

# save to directory

```{r}
ggsave(filename = "plots/survival_TCGA_BLCA.pdf",
       plot = print(p, newpage = F),
       width = 3,
       height = 3)
```
