---
title: "Lesson 5a Challenge problems"
author: "Magdalena Russell"
date: '2022-06-15'
output: html_document
---

All exercises use the clinical object we created in the readings:

```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)

# Load the data
clinical <- read.csv("data/clinical.csv")
```

Create a new object from clinical called demographics_BLCA that includes only the height, bmi, gender, ethnicity, and race.

```{r}
# your code here...

```

Code as many ways to select the columns year_of_diagnosis, paper_Tobacco.smoking.history, pack_years_smoked, cigarettes_per_day from the clinical object

```{r}
# your code here...

```

Create a new object from clinical called demographics_BLCA_current_smokers that includes only the height, bmi, gender, ethnicity, and race. However, restrict data such that only current smokers are included (hint: relevant column is called Tobacco.smoking.history).

```{r}
# your code here...

```


