---
title: "Lesson 2b In-class exercises"
author: "Magdalena Russell"
date: '2022-06-15'
output: html_document
---

__Under each heading below, type your code and run it. Feel free to add notes and comments to help you learn.__ 


# Using a function: Rounding numbers

```{r}
## [Round 3.14 using the round() function]

```

What is the answer?


# Using a function with more arguments

```{r}
## [add the digits parameter to your rounding code above and round to 1 digit]


```


# Getting help with the `round()` function

```{r}
?round
#run me
```

What happened when you ran the above code chunk?



# Using round with complete argument labels

```{r}
# [fill in the values we used above after the equals]
round(x = , digits = ) # edit me
```

Did you get the same answer as before? Or different?



# Switching the order of arguments

```{r}
# [copy your code from the last code chunk and switch the positions of the parameters, such that x = 3.14 is after the parenthesis]

```

What happened?



# Removing argument labels

```{r}
# [copy your round statement from above and delete the labels]

```

Now what happened? Did you get the same answer as before?
